### 2019-10-31 analysis_and_paper/prediction_plots.Rmd

```{r}
library("msm")
library("spatialfil")
```

```{r prediction-setup, cache = TRUE}
sd0 <- 0.6
kern <- convKernel(sigma = 8, k = "gaussian")

nrmlz <- function(mat) {
	tot <- sum(mat)
	mat/tot
}

thetaCons <- seq(-1.5, 1.5, .02)
thetaLib <- seq(-1.5, 1.5, .02)

combiThetas <- expand.grid(
  Liberals = thetaLib
  , Conservatives = thetaCons
  , figType = c("Model", "Prediction")
  , Model = c("No effect", "One mind", "Two minds", "Any effect")
)
combiThetas$Density <- NA

combiThetasRaw <- expand.grid(
  Liberals = thetaLib
  , Conservatives = thetaCons
)

makeTeamBPlot <- function(z, dat, ylab, xlab, breaks, strip = FALSE){
  graycols <- gray.colors(100, max(.0, .6 - 100 * max(dat$Density))
                          , max(.7, 1 - 50 * max(dat$Density)))
  ggplot(
    mapping = aes(y = Liberals, x = Conservatives, z = eval(parse(text = z)), fill = ..level..)
    , data = dat
  ) +
  stat_contour(geom = "polygon", bins = 20) +
  geom_hline(yintercept = 0, color = "grey70", size = .3, alpha = .3) +
  geom_vline(xintercept = 0, color = "grey70", size = .3, alpha = .3) +
  scale_fill_gradientn(colors = rev(graycols)) +
  xlim(c(min(combiThetas$Conservatives), max(combiThetas$Conservatives))) +
  ylim(c(min(combiThetas$Liberals), max(combiThetas$Liberals))) +
  coord_fixed() +
  labs(x = xlab, y = ylab) +
  facet_grid(Model ~ figType) + 
  theme(
    legend.position = "none"
    , axis.title.x = if(xlab == "") element_blank() else element_text()
    , axis.text.x = if(xlab == "") element_blank() else element_text()
    , axis.ticks.x = if(xlab == "") element_blank() else element_line()
    , strip.text.x = if(!strip) element_blank() else element_text()
    , panel.border = element_rect(color = "black", fill = NA)
    , panel.spacing = unit(rel(10), "points")
    , strip.text.y = ggplot2::element_text(margin = ggplot2::margin(0, 0, 0, ggplot2::rel(12)))
  )
}
```

```{r null-model-predicitons, warning = FALSE}
nullTruth <- function(theta, sd){
  dnorm(theta[1], 0, sd) * dnorm(theta[2], 0, sd)
}

modelDensityNull <- apply(combiThetasRaw
                                              , 1
                                              , nullTruth
                                              , sd = .01)

modelDensityNull <- nrmlz(modelDensityNull)

mat <- matrix(modelDensityNull, nrow = length(thetaLib), ncol = length(thetaCons))
predmat <- nrmlz(applyFilter(mat, kern))
predictionDensityNull <- as.vector(predmat) * 100

combiThetas[combiThetas$figType == "Model" & combiThetas$Model == "No effect", "Density"] <- modelDensityNull
combiThetas[combiThetas$figType == "Prediction" & combiThetas$Model == "No effect", "Density"] <- predictionDensityNull

nullmod <- data.frame(x = 0, y = 0, figType = "Model")
```

```{r one-mind-predictions, cache = TRUE, dependson = "prediction-setup"}
#Prior Model 
same_direction <- function(theta, Sigma){
  dtnorm(theta[2], 0, Sigma, lower = 0) * dtnorm(theta[1], 0, Sigma, lower = 0)}

model_same_direction  <- apply(combiThetasRaw
                                                    , 1
                                                    , same_direction
                                                    , Sigma = sd0)
model_same_direction <- nrmlz(model_same_direction)

mat <- matrix(model_same_direction
              , nrow = length(thetaCons)
              , ncol = length(thetaCons))
predmat <- nrmlz(applyFilter(mat, kern))
prediction_same_direction <- as.vector(predmat)

combiThetas[combiThetas$figType == "Model" & combiThetas$Model == "One mind", "Density"] <- model_same_direction
combiThetas[combiThetas$figType == "Prediction" & combiThetas$Model == "One mind", "Density"] <- prediction_same_direction
```

```{r two-mind-predictions, cache = TRUE, dependson = "prediction-setup"}
#Prior Model 
opposite_direction <- function(theta, Sigma){
  dtnorm(theta[2], 0, Sigma, lower = 0) * dtnorm(theta[1], 0, Sigma, upper = 0)}

model_opposite_direction  <- apply(combiThetasRaw
                                                    , 1
                                                    , opposite_direction
                                                    , Sigma = sd0)
model_opposite_direction <- nrmlz(model_opposite_direction)

mat <- matrix(model_opposite_direction
              , nrow = length(thetaCons)
              , ncol = length(thetaCons))
predmat <- nrmlz(applyFilter(mat, kern))
prediction_opposite_direction <- as.vector(predmat)

combiThetas[combiThetas$figType == "Model" & combiThetas$Model == "Two minds", "Density"] <- model_opposite_direction
combiThetas[combiThetas$figType == "Prediction" & combiThetas$Model == "Two minds", "Density"] <- prediction_opposite_direction
```

```{r unconstrained-predictions, cache = TRUE, dependson = "prediction-setup"}
# Prior Model
genTruth <- function(theta, sd){
  dnorm(theta[1], 0, sd) * dnorm(theta[2], 0, sd)
}

modelDensityGeneral <- apply(combiThetasRaw
                                              , 1
                                              , genTruth
                                              , sd = sd0)
modelDensityGeneral <- nrmlz(modelDensityGeneral)

# Model Predictions
mat <- matrix(modelDensityGeneral
              , nrow = length(thetaCons)
              , ncol = length(thetaCons))
predmat <- nrmlz(applyFilter(mat, kern))
predictionDensityGeneral <- as.vector(predmat)

combiThetas[combiThetas$figType == "Model" & combiThetas$Model == "Any effect", "Density"] <- modelDensityGeneral
combiThetas[combiThetas$figType == "Prediction" & combiThetas$Model == "Any effect", "Density"] <- predictionDensityGeneral
```

```{r model-plots, fig.width = 4.5, fig.height = 5.4, eval = FALSE}
# plot_labels <- data.frame(
#   Liberals = 1.25
#   , Conservatives = -0.725
#   , Model = c("One mind", "Two minds")
#   , label = c("One mind", "Two minds")
# )

otm1_models <- subset(combiThetas, figType == "Model") %>%
  mutate(Model = factor(Model, levels = c("One mind", "Two minds", "No effect", "Any effect"))) %>%
  group_by(Model) %>%
  do(mutate(., Density = exp(log(Density) - max(log(Density))))) %>%
  ggplot(
    mapping = aes(
      y = Liberals
      , x = Conservatives
    )
  ) +
  stat_contour(geom = "polygon", aes(z = Density, fill = ..level..)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  # scale_fill_gradient(low = bg_col, high = "#385E49") +
  geom_point(data = data.frame(Liberals = 0, Conservatives = 0, Model = "No effect"), color = "#385E49", size = 3) +
  # geom_text(data = plot_labels, aes(label = label)) +
  coord_fixed() +
  labs(
    x = expression(alpha)
    , y = expression(beta)
  ) +
  facet_wrap(~ Model, ncol = 2) +
  theme(
    legend.position = "none"
    , axis.text.x = element_blank()
    , axis.text.y = element_blank()
    , axis.ticks.x = element_blank()
    , axis.ticks.y = element_blank()
    , panel.border = element_rect(color = "black", fill = NA)
    , panel.spacing = unit(rel(10), "points")
    , strip.text.y = ggplot2::element_text(margin = ggplot2::margin(0, 0, 0, ggplot2::rel(12)))
  )
```

```{r prediction-plots}
otm1_prediction_plot <- subset(combiThetas, figType == "Prediction") %>%
  mutate(Model = factor(Model, levels = c("One mind", "Two minds", "No effect", "Any effect"))) %>%
  group_by(Model) %>%
  do(mutate(., Density = exp(log(Density) - max(log(Density))))) %>%
  ggplot(
    plot_data
    , mapping = aes(
      y = Liberals
      , x = Conservatives
      , z = Density
      , fill = ..level..
    )
  ) +
  stat_contour(geom = "polygon") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  scale_fill_gradient(low = "white", high = "grey30") +
  coord_fixed() +
  labs(
    x = expression("Rating score difference"~group("[", Delta~italic(z)~score, "]"))
    , y = expression("IAT score difference"~group("[", Delta~italic(z)~score, "]"))
  ) +
  facet_wrap(~ Model, ncol = 2) +
  theme(
    legend.position = "none"
    , axis.text.x = element_blank()
    , axis.text.y = element_blank()
    , axis.ticks.x = element_blank()
    , axis.ticks.y = element_blank()
    , panel.border = element_rect(color = "black", fill = NA)
    , panel.spacing = unit(rel(10), "points")
    , plot.title = element_text(hjust = 0.5)
    , strip.text.x = element_text(margin = margin(b = rel(8)))
    , strip.text.y = element_text(margin = margin(0, 0, 0, rel(12)))
  )
```