### 2019-10-31 otm2pilot2/results/analysis2pilot2_florida.Rmd

---
title             : "Analysis of OTM 2 pilot 2: Florida"
author            : "Frederik Aust & Tobias Heycke"
date              : "`r format(Sys.time(), '%d %B, %Y')`"

output:
  bookdown::html_document2:
    theme         : "spacelab"
    df_print      : "kable"
    code_folding  : "hide"
    toc           : true
    toc_float     : true
---

```{r init, include = FALSE}
library("magrittr")
library("tidyr")
library("dplyr")
library("assertthat")

library("papaja")
library("ggplot2")
library("ggforce")
library("cowplot")

library("afex")
library("emmeans")
library("BayesFactor")

source("https://gist.githubusercontent.com/crsh/bd4d1f62d300462ea0c0f44b9ad38616/raw/edd9c74e24b68f42433c2526cafc888509b1b8bc/batch_download_github.R")
source("https://gist.githubusercontent.com/crsh/357458c41fd3d554fb24/raw/f7725d5c4894a055a1b2e461dc5c39f3db23b2b8/batch_read.R")

source("../../analysis_and_paper/modal_frame_rate.R")

# Set seed for random number generator
set.seed(315054738)

# Use effect coding
options(contrasts = c("contr.sum", "contr.poly"))

# Use multivariate models for emmeans contrasts and post-hoc tests
afex_options(emmeans_model = "multivariate")

# Configure df approximation for mixed model contrasts and post-hoc tests
emm_options(
  lmer.df = "satterthwaite"
  , lmerTest.limit = 22384
)

#t test print function
printBFt <- function(BF, HStyle = 0, index = 1, OutputSize = 99999.99 , postit = 1000000, N = FALSE){
  if(as.vector(BF[index]) < 1 & HStyle == 0){
    b <- 1/as.vector(BF[index])
    num <- "01"
  }else{
    b <- as.vector(BF[index])
    num <- "10"
  }
  s <- " "
  if(N == TRUE) s <- paste0('(', nrow(BF@data), ') ') 
  if(as.character(class(BF@numerator[[names(BF@numerator)[index]]])) == "BFoneSample"){
    rBF <- BayesFactor::ttestBF(BF@data[,1], mu = BF@numerator[[names(BF@numerator)[index]]]@prior$mu, rscale = BF@numerator[[names(BF@numerator)[index]]]@prior$rscale)
  }
  if(as.character(class(BF@numerator[[names(BF@numerator)[1]]])) == "BFindepSample"){
      rBF <- BayesFactor::ttestBF(subset(BF@data, BF@data[,2] == "x")[,1] , subset(BF@data, BF@data[,2] == "y")[,1], rscale = BF@numerator[[names(BF@numerator)[index]]]@prior$rscale, paired = FALSE)
    }
  post <- BayesFactor::posterior(rBF, index = index, iterations = postit)
  d <- median(post[, "delta"])
  HDI <- coda::HPDinterval(post[, "delta"])
  ifelse(b > OutputSize, paste0('$\\mathit{BF}_{', num, '}', s, '= ', as.scientific(b), '$', ', ', '$d = ', printnum(d), '$', ', ', '95% HDI [', printnum(HDI[1]), ', ', printnum(HDI[2]), ']'), paste0('$\\mathit{BF}_{', num, '}', s, '= ', printnum(b), '$', ', ', '$d = ', printnum(d), '$', ', ', '95% HDI [', printnum(HDI[1]), ', ', printnum(HDI[2]), ']'))
}
```

```{r otm2pilot2f-analysis-preferences, cache = TRUE}
# Data wrangling
process_rawdata <- TRUE

# Data location
raw_data_path <- "data_raw_florida/"
processed_data_path <- "data_processed_florida/"

# Default ggplot theme
theme_set(theme_apa())

## Ugly hack to overcome bug in ggplot2, https://github.com/tidyverse/ggplot2/issues/2058
assignInNamespace("theme_nothing", function() {
    theme_void() + theme(axis.text.x = NULL, axis.text.y = NULL, axis.line.x = element_blank(), axis.line.y = element_blank())
}, "cowplot")

# Number of MCMC samples for Bayesian analysis
otm2_n_mcmc_samples <- 1e6
```

```{r otm2pilot2f-load-data, cache = TRUE, dependson = "otm2pilot2f-analysis-preferences", warning = FALSE}
# Process raw data
if(process_rawdata) {

  # Merge raw data
  otm2p2f_eval <- batch_read(
    raw_data_path
    , pattern = "Eval"
    , recursive = TRUE
    , read_fun = read.delim
  )

  otm2p2f_mem <- batch_read(
    raw_data_path
    , pattern = "MemTest"
    , recursive = TRUE
    , read_fun = read.delim
  )

  otm2p2f_log <- batch_read(
    raw_data_path
    , pattern = "ScreenLog"
    , recursive = TRUE
    , read_fun = read.csv
    , header = FALSE
  )

  otm2p2f_analysis_factors <- c("ParticipantNumber", "Location", "Block", "ValenceBlock")

  otm2p2f_mem$ParticipantNumber <- as.integer(otm2p2f_mem$ParticipantNumber)
  #otm2p2_mem <- left_join(otm2p2_mem, otm2p2_demo, by = "ParticipantNumber")

  otm2p2f_mem <- otm2p2f_mem %>%
    mutate(
      ParticipantNumber = factor(ParticipantNumber)
      , ValenceBlock = ifelse(ValenceBlock == 1, "negPrime - posPrime", "posPrime - negPrime") %>% factor
      , Block = factor(Block)
      , Accuracy = as.integer(NumbercorrectIdent) / 20
    )

    # Save processed data
  saveRDS(otm2p2f_eval, paste0(processed_data_path, "otm2p2f_eval.rds"))
  saveRDS(otm2p2f_mem, paste0(processed_data_path, "otm2p2f_memory.rds"))
} else {
  otm2p2f_eval <- readRDS(paste0(processed_data_path, "otm2p2f_eval.rds"))
  otm2p2f_mem <- readRDS(paste0(processed_data_path, "otm2p2f_memory.rds"))
}
```

```{r otm2p2f-recognition, results = "asis"}
otm2p2f_mem_ttest <- t.test(otm2p2f_mem$Accuracy, mu = 0.5
                          , alternative = "greater") %>%
  apa_print %$%
  full_result

otm2p2f_mem_ttestBF <- ttestBF(otm2p2f_mem$Accuracy - 0.5
                             , mu = 0, rscale = "medium"
                             , nullInterval = c(0, Inf))[1] %>%
  printBFt

```

# Results

The identification performance was not significantly better than chance, `r otm2p2f_mem_ttest`, however the Bayes Factor was inconclusive `r otm2p2f_mem_ttestBF`.

```{r otm2p2f-recognition-viz, fig.width = 3, fig.height = 4, fig.align = "center"}
ggplot(otm2p2f_mem, aes(x = 1, y = Accuracy)) +
  geom_violin() +
  geom_hline(yintercept = 0.5, color = grey(0.7), linetype = "22") +
  geom_sina() +
  stat_summary(fun.data = mean_cl_boot) +
  theme(
    axis.text.x = element_blank()
    , axis.title.x = element_blank()
    , axis.ticks.x = element_blank()
  )
```