### 2021-09-10 otm2/preregistration/power_analyses.Rmd
Preparations for Experiment 2 data collection

---
title             : "Power analyses for OTM2"
author            : "Frederik Aust"
date              : "`r format(Sys.time(), '%d %B, %Y')`"

output:
  bookdown::html_document2:
    theme         : "spacelab"
    df_print      : "kable"
    code_folding  : "hide"
    toc           : true
    toc_float     : true
---

```{r setup, include = FALSE}
library("tidyr")
library("dplyr")
library("ggplot2")
library("cowplot")
library("papaja")

library("pwr")
library("emmeans")

# devtools::install_github("arcaldwell49/Superpower")
library("Superpower")

source("../../analysis_and_paper/effect_sizes.R")
```

Because it grants greater flexibility for ANOVA designs, I here used the in-development R package `Superpower`.
The make sure I'm doing this right, I first reproduced the power analyses done in G*Power for OTM1.

# OTM 1

The following gives the interaction power for the effect size estimated in the G*Power sensitivity analysis

```{r}
contrast_code <- tibble::tibble(
  a = rep(contr.sum(2), 2)
  , b = rep(contr.sum(2), each = 2)
  , x = a*b * -1
)$x

f <- 0.1481313
mu <- f * contrast_code
n <- 75
r <- 0.5
alpha_level <- 0.05

string <- "2w*2b"
labelnames <- c("block", "1", "2", "order", "n-p", "p-n")

design_result <- ANOVA_design(
  design = string
  , n = n
  , mu = mu
  , sd = 1
  , r = r
  , labelnames = labelnames
)

otm1_power <- ANOVA_exact(
  design_result
  , alpha_level = alpha_level
  , emm = TRUE
  , verbose = FALSE
  , contrast_type = "pairwise"
  , emm_comp = "block | order"
)

otm1_power$main_results
```

as well as the power for the block contrasts

```{r}
otm1_power$emm_results
```

However, this doesn't give the result for the desired contrast that collapses over the counterbalancing factor, nor for a directional test.
I contributed a pull request to `Superpower` to enable power analyses for contrasts specified with `emmeans`.
The following gives the result for the collapsed contrast

```{r}
custom_contrast <- contrast(
  emmeans(
    otm1_power$emmeans$emmeans
    , specs = ~ order * block
  )
  , list(block_effect = c(0.5, -0.5, -0.5, 0.5))
)

emmeans_power(custom_contrast, alpha_level = 0.05)
```

which is the same as doing a within-subject t-test in this case.

```{r}
pwr.t.test(n = 150, d = 2*f, sig.level = 0.05, type = "paired")
```

To get the power for the directional test, I can increase the alpha level as needed, which shows that this contrast is the same as the within-subject t-test collapsing data across groups.

```{r}
emmeans_power(custom_contrast, alpha_level = 0.1)
pwr.t.test(150, sig.level = 0.05, d = 2*f, type = "paired", alternative = "greater")
```

This now also allows me to get the power for the one-sided tests of the uncollapsed contrasts

```{r}
emmeans_power(otm1_power$emmeans$contrasts, alpha_level = 0.1)
```

Anyway, the following reproduces the G*Power sensitivity analysis for the collapsed contrast.

```{r}
pwr.t.test(150, sig.level = 0.05, power = 0.95, type = "paired", alternative = "greater")
```

<!--

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

## More elaborate considerations of repeated-measure correlations

The following is a table of repeated measure correlations from OTM1 across locations.

```{r}
readRDS("../../otm1/results/data_processed/otm1_attitudes_cleaned.rds") %>%
  mutate(
      Eval = scale(Eval) %>% as.numeric
      , IATscore = scale(IATscore) %>% as.numeric
  ) %>%
  gather("Measure", "Attitude", Eval, IATscore) %>%
  mutate(Measure = factor(Measure)) %>%
  pivot_wider(names_from = c("Block", "Measure"), values_from = "Attitude") %>%
  select_if(is.numeric) %>%
  psych::corr.test()
```

```{r}
otm1_power <- function(f, rm_cor, n = 76, alpha = 0.05) {
  contrast_code <- tibble::tibble(
    a = rep(contr.sum(2), 2)
    , b = rep(contr.sum(2), each = 2)
    , x = a*b * -1
  )$x

  mu <- f * contrast_code

  cor_mat <- diag(4)
  triangular <- c(0, 1, 0, 0, 1, 0)
  cor_mat[upper.tri(cor_mat)] <- triangular * rm_cor
  cor_mat[lower.tri(cor_mat)] <- triangular * rm_cor

  string <- "2w*2b"
  labelnames <- c("block", "1", "2", "order", "n-p", "p-n")

  design_result <- ANOVA_design(
    design = string
    , n = n
    , mu = mu
    , sd = 1
    , r = cor_mat
    , labelnames = labelnames
  )

  otm1_power <- ANOVA_exact(
    design_result
    , alpha_level = alpha
    , emm = TRUE
    , verbose = FALSE
    , contrast_type = "pairwise"
    , emm_comp = "block | order"
  )

  otm1_power
}
```

### Evaluative ratings

```{r}
eval_rm_cor <- -0.8
otm1_f_eval_anova <- 0.27918

otm1_power_eval_anova <- otm1_power(f = otm1_f_eval_anova, rm_cor = eval_rm_cor)
otm1_power_eval_anova_int <- otm1_power_eval_anova$main_result

otm1_power_eval_anova_int$generalized_eta_squared <- afex::aov_4(
  y ~ block * order + (block | subject)
  , data = otm1_power_eval_anova$dataframe
)$anova_table$ges
otm1_power_eval_anova_int
```

Without Bonferroni-correction for the order-wise two-sided contrasts

```{r}
otm1_f_eval_contrast <- 0.39482
emmeans_power(otm1_power(f = otm1_f_eval_contrast, rm_cor = eval_rm_cor)$emmeans$contrasts, alpha_level = 0.05)
```

For the collapsed two-sided contrasts

```{r}
otm1_power(f = 0.2543, rm_cor = eval_rm_cor)$emmeans$emmeans %>%
  emmeans(specs = ~ block * order) %>%
  contrast(method = list(collapsed = c(-0.5, 0.5, 0.5, -0.5))) %>% 
  emmeans_power(alpha_level = 0.1)
```

### IAT scores

```{r}
iat_rm_cor <- -0.6 * sqrt(0.5^2)
otm1_f_iat_anova <- 0.23726

otm1_power_iat_anova <- otm1_power(f = otm1_f_iat_anova, rm_cor = iat_rm_cor)
otm1_power_iat_anova_int <- otm1_power_iat_anova$main_result

otm1_power_iat_anova_int$generalized_eta_squared <- afex::aov_4(
  y ~ block * order + (block | subject)
  , data = otm1_power_iat_anova$dataframe
)$anova_table$ges
otm1_power_iat_anova_int
```

Without Bonferroni-correction for the order-wise two-sided contrasts

```{r}
otm1_f_iat_contrast <- 0.335535
emmeans_power(otm1_power(f = otm1_f_iat_contrast, rm_cor = iat_rm_cor)$emmeans$contrasts, alpha_level = 0.05)
```

For the collapsed one-sided contrasts

```{r}
otm1_power(f = 0.21611, rm_cor = iat_rm_cor)$emmeans$emmeans %>%
  emmeans(specs = ~ block * order) %>%
  contrast(method = list(collapsed = c(-0.5, 0.5, 0.5, -0.5))) %>% 
  emmeans_power(alpha_level = 0.1)
```

```{r eval = TRUE}
knitr::opts_chunk$set(eval = TRUE)
```

## Simplified again

-->

## With implicit repeated measures correlation

```{r}
rydell_pes <- 0.1
rydell_f <- 4.78
rydell_n <- 50
rydell_dz <- pes_to_dz(rydell_pes, rydell_n, n_groups = 2)

heycke_t <- c(2.54, 1.45)
heycke_n <- c(28, 26)
heycke_dz <- heycke_t / sqrt(heycke_n)
heycke_pes <- dz_to_pes(heycke_dz, heycke_n, n_groups = 1)
# heycke_pes <- (heycke_dz^2 * heycke_n) / (heycke_dz^2 * heycke_n + heycke_n - 1) # pes

# mean(c(heycke_dz, rydell_dz))
# mean(c(heycke_pes, rydell_pes))
```

To simplify this, I will focus on the power for what we'd expect to be the smallest effect: The *learning block* contrast within each *valence order* group in IAT scores.
Rydell et al. report the smallest contrast as $\eta_p^2 = `r rydell_pes`$ ($d_z \approx `r rydell_dz`$), which is comparable in magnitude (but not direction) to the corresponding effect sizes reported by Heycke et al.

```{r}
otm1_power <- function(f, rm_cor, n = 76, alpha = 0.05) {
  contrast_code <- tibble::tibble(
    a = rep(contr.sum(2), 2)
    , b = rep(contr.sum(2), each = 2)
    , x = a*b * -1
  )

  mu <- with(
    contrast_code
    , f * x +
      - 0.25 * a
  )

  cor_mat <- diag(4)
  triangular <- c(0, 1, 0, 0, 1, 0)
  cor_mat[upper.tri(cor_mat)] <- triangular * rm_cor
  cor_mat[lower.tri(cor_mat)] <- triangular * rm_cor

  string <- "2w*2b"
  labelnames <- c("block", "1", "2", "order", "n-p", "p-n")

  design_result <- ANOVA_design(
    design = string
    , n = n
    , mu = mu
    , sd = 1
    , r = cor_mat
    , labelnames = labelnames
  )

  otm1_power <- ANOVA_exact(
    design_result
    , alpha_level = alpha
    , emm = TRUE
    , verbose = FALSE
    , contrast_type = "pairwise"
    , emm_comp = "block | order"
  )

  otm1_power
}
```

```{r}
otm1_n <- 76

# 95% power
otm1_pes <- 0.08068
otm1_dz <- pes_to_dz(otm1_pes, otm1_n, n_groups = 2)
# rho <- 0.5
# f <- dz_to_f(otm1_dz, rho)
# 
# otm1_contrast_power <- otm1_power(f = f, n = otm1_n, rm_cor = rho)$emmeans$contrasts %>%
#   emmeans_power(alpha = 0.05) %>%
#   .[1, ]
# otm1_contrast_power

# 80% power
otm1_pes2 <- 0.050336
otm1_dz2 <- pes_to_dz(otm1_pes2, otm1_n, n_groups = 2)
# f <- dz_to_f(otm1_dz2, rho)
# 
# otm1_contrast_power2 <- otm1_power(f = f, n = otm1_n, rm_cor = rho)$emmeans$contrasts %>%
#   emmeans_power(alpha = 0.05) %>%
#   .[1, ]
# otm1_contrast_power2
```

Across all labs, our experimental design has 95% power to detect effects as small as $\eta^2_p = `r printp(otm1_pes)`$ ($\delta_z = `r otm1_dz`$) and 80% power to detect effects as small as $\eta^2_p = `r printp(otm1_pes2)`$ ($\delta_z = `r otm1_dz2`$).
Figure\ \@ref(fig:pes-plot) illustrates the implied Cohen's $\delta$ depending on the assumed repeated-measures correlation $\rho$. 
That is, we have sufficient power to find (or rule out) effects approximately 20% smaller than the typical effects reported for the weakest contrast.

```{r pes-plot, fig.width = 5, fig.height = 4, fig.cap = "Sensitivity curve for OTM1"}
tibble(
  rho = rep(seq(-1, 1, length.out = 100), 2)
  , delta = c(
    dz_to_f(otm1_dz, unique(rho)) * 2
    , dz_to_f(otm1_dz2, unique(rho)) * 2
  )
  , power = rep(c("0.95", "0.80"), each = length(rho)/2)
) %>% 
  mutate(power = relevel(factor(power), "0.95")) %>% 
  ggplot(aes(x = rho, y = delta, color = power, group = power)) +
  geom_line() +
  scale_color_viridis_d(end = 0.8) +
  labs(
    x = bquote("Correlation between learning blocks ["*rho*"]")
    , y = bquote("Cohen's"~delta)
    , color = bquote("Power"~"["*1 - beta*"]")
  ) +
  theme_apa(box = TRUE) + 
  theme(
    panel.grid.major.x = element_line()
    , panel.grid.major.y = element_line()
    , panel.grid.minor.x = element_line()
    , panel.grid.minor.y = element_line()
    , legend.position = c(0.99, 0.99)
    , legend.justification = c(1, 1)
    , legend.background = element_rect(fill = scales::alpha("white", 0.5))
  )
```

# OTM 2

<!--

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

The following gives the senseitivity analyses for the three-way interaction.
Because `superpower` only allows for power analyses, I manually estimated the effect size that yields approximately 95% power.

```{r}
contrast_codes <- tibble::tibble(
  a = rep(contr.sum(2), 4)
  , b = rep(rep(contr.sum(2), each = 2), 2)
  , c = rep(contr.sum(2), each = 4)
  , x2 = a*c * -1
  , x3 = a*b*c * -1
)

f <- 0.0807625
mu <- contrast_codes$x3 * f
n <- 125
sd <- 1
r <- 0.5
alpha_level <- 0.05

string = "2w*2b*2b"
labelnames <- c("block", "1", "2", "order", "p-n", "n-p", "time", "13", "20")

design_result <- ANOVA_design(
  design = string
  , n = n
  , mu = mu
  , sd = 1
  , r = r
  , labelnames = labelnames
)

otm2_power <- ANOVA_exact(
  design_result
  , alpha_level = alpha_level
  , emm = TRUE
  , verbose = FALSE
  , contrast_type = "pairwise"
  , emm_comp = "block | time * order"
)

otm2_power$main_results
```

Thus, we have 95% power to detect a three-way interaction between *order*, *block*, and *presentation time* of $f = `r round(f, 2)`$ ($d = `r round(2*f, 2)`$).

The two-way interaction contrasts the same number of cells and participants and, thus, we have the same power here.

```{r}
f <- 0.0807625
mu <- contrast_codes$x2 * f

design_result <- ANOVA_design(
  design = string
  , n = n
  , mu = mu
  , sd = 1
  , r = r
  , labelnames = labelnames
)

otm2_power <- ANOVA_exact(
  design_result
  , alpha_level = alpha_level
  , emm = TRUE
  , verbose = FALSE
  , contrast_type = "pairwise"
  , emm_comp = "block | time"
)

otm2_power$main_results
```

For example, we have 95% power to detect a two-way interaction between *block*, and *presentation time* of $f = `r round(f, 2)`$ ($d = `r round(2*f, 2)`$).

If we want to test the *block* effects (collapsed across *order*) separately for each presentation time, we get different sensitivities depending on whether we perform correction for multiple comparisons.

```{r}
f <- 0.104169
mu <- contrast_codes$x2 * f

design_result <- ANOVA_design(
  design = string
  , n = n
  , mu = mu
  , sd = 1
  , r = r
  , labelnames = labelnames
)

otm2_power <- ANOVA_exact(
  design_result
  , alpha_level = alpha_level
  , emm = TRUE
  , verbose = FALSE
  , contrast_type = "pairwise"
  , emm_comp = "block | time"
)

emmeans_power(otm2_power$emmeans$contrasts, alpha_level = 0.1)
```

Without correction, we have 95% power to detect a block effect within *presentation times* of $f = `r round(f, 2)`$ ($d = `r round(2*f, 2)`$).

```{r}
f <- 0.1142145
mu <- contrast_codes$x2 * f

design_result <- ANOVA_design(
  design = string
  , n = n
  , mu = mu
  , sd = 1
  , r = r
  , labelnames = labelnames
)

otm2_power <- ANOVA_exact(
  design_result
  , alpha_level = alpha_level
  , emm = TRUE
  , verbose = FALSE
  , contrast_type = "pairwise"
  , emm_comp = "block | time"
)

emmeans_power(otm2_power$emmeans$contrasts, alpha_level = 0.05)
```

With bonferroni correction, we have 95% power to detect a block effect within *presentation times* of $f = `r round(f, 2)`$ ($d = `r round(2*f, 2)`$).

In case presentation time has no effect, we can collapse across presentation durations.

```{r}
f <- 0.0736605
mu <- contrast_codes$x2 * f

design_result <- ANOVA_design(
  design = string
  , n = n
  , mu = mu
  , sd = 1
  , r = r
  , labelnames = labelnames
)

otm2_power <- ANOVA_exact(
  design_result
  , alpha_level = alpha_level
  , emm = TRUE
  , verbose = FALSE
  , contrast_type = "pairwise"
  , emm_comp = "block | time"
)

custom_contrast <- contrast(
  emmeans(
      otm2_power$emmeans$emmeans
      , specs = ~ time * block
  )
  , list(block_effect = c(-0.5, 0.5, 0.5, -0.5))
)

emmeans_power(custom_contrast, alpha_level = 0.1)
```

Here, we have 95% power to detect a block effect of $f = `r round(f, 2)`$ ($d = `r round(2*f, 2)`$).

## Time x Measure x Block

Because `superpower` only allows for power analyses, I manually determined the sample size that yields approximately 95% power.
The following assumes that we are not interested in the effect of the counterbalancing factor *order*.

```{r fig.width = 7, fig.height = 7}
contrast_codes <- tibble::tibble(
  time = rep(contr.sum(2), 4)
  , block = rep(rep(contr.sum(2), each = 2), 2)
  , measure = rep(contr.sum(2), each = 4)
  , time_block = time * block
  , time_measure = time * measure
  , block_measure = block * measure
  , time_block_measure = time * block * measure
)

f <- 0.5
mu <- with(
  contrast_codes
  , block * f +
    block_measure * f + 
    -time_block * 0.25*f + 
    time_block_measure * 0.25*f
)
n <- 68
# r <- 0.5
alpha_level <- 0.05

eval_reliability <- 0.8
iat_reliability <- 0.5
measure_cor <- 0.8 * sqrt(eval_reliability * iat_reliability)

string <- "2w*2w*2b"
labelnames <- c("measure", "eval", "iat", "block", "1", "2", "time", "13", "20")

eval_cor_mat <- iat_cor_mat <- inter_cor_mat <- diag(4)
triangular <- c(0, 1, 0, 0, 1, 0)

eval_cor_mat[upper.tri(eval_cor_mat)] <- triangular * -eval_reliability
eval_cor_mat[lower.tri(eval_cor_mat)] <- triangular * -eval_reliability
iat_cor_mat[upper.tri(eval_cor_mat)] <- triangular * -iat_reliability
iat_cor_mat[lower.tri(eval_cor_mat)] <- triangular * -iat_reliability

inter_cor_mat <- inter_cor_mat * measure_cor
inter_cor_mat[upper.tri(eval_cor_mat)] <- triangular * -measure_cor
inter_cor_mat[lower.tri(eval_cor_mat)] <- triangular * -measure_cor

cor_mat <- rbind(cbind(eval_cor_mat, inter_cor_mat), cbind(inter_cor_mat, iat_cor_mat))

design_result <- ANOVA_design(
  design = string
  , n = n
  , mu = mu
  , sd = 1
  , r = 0.5
  , labelnames = labelnames
)
dimnames(cor_mat) <- dimnames(design_result$cor_mat)

design_result <- ANOVA_design(
  design = string
  , n = n
  , mu = mu
  , sd = 1
  , r = cor_mat
  , labelnames = labelnames
)

otm2_power <- ANOVA_exact(
  design_result
  , alpha_level = alpha_level
  , emm = TRUE
  , verbose = FALSE
  , contrast_type = "pairwise"
  , emm_comp = "block | time * measure"
)

otm1_observed_plot <- readRDS("../../otm1/results/data_processed/otm1_attitudes_cleaned.rds") %>%
    mutate(
      Block = ifelse(ValenceBlock == "Negative-positive", 3 - as.numeric(as.character(Block)), Block) %>% factor
    ) %>%
    select(-ValenceBlock) %>%
    mutate(Rating = scale(Eval) %>% as.vector, `IAT score` = scale(IATscore) %>% as.vector) %>%
    gather("Measure", "Attitude", Rating, `IAT score`) %>% 
    mutate(Measure = factor(Measure)) %>%
    ggplot(aes(x = Block, y = Attitude)) +
      stat_summary(fun.data = mean_cl_normal) +
      facet_grid("" ~ Measure) +
      coord_cartesian(ylim = c(-1.25, 1.25)) +
      theme_bw() +
      ggtitle("Results of Experiment 1")

otm2_expected_plot <- otm2_power$dataframe %>% 
  mutate(
    measure = ifelse(measure == "measure_eval", "Rating", "IAT score")
    , block = ifelse(block == "block_1", "1", "2")
    , time = ifelse(time == "time_13", "13 ms", "20 ms")
  ) %>%
  ggplot(aes(x = block, y = y)) +
    stat_summary(fun.data = mean_cl_normal) +
    facet_grid(time ~ measure) +
    coord_cartesian(ylim = c(-1.25, 1.25)) +
    theme_bw() +
    ggtitle("Assumed for power analysis") +
    labs(x = "Block", y = "Attitude")

plot_grid(
  otm1_observed_plot
  , otm2_expected_plot
  , ncol = 1
  , rel_heights = c(1, 1.5)
)

otm2_power$main_results
```

Thus, we have 95% power to detect a three-way interaction between *measure*, *block*, and *presentation time* of $f = `r round(0.25*f, 2)`$ ($d = `r round(2*0.25*f, 2)`$) and more than 95% power to detect a two-way interaction between *measure*, *block* of $f = `r round(f, 2)`$ ($d = `r round(f, 2)`$) for $n = `r n`$ ($N = `r 2*n`$).
These results are, of course, contigent on the assumed reliabilities of the rating measure ($r = `r #round(eval_reliability, 2)`$), the IAT score ($r = `r #round(iat_reliability, 2)`$), and the correlation among the two measures ($r = `r #round(measure_cor, 2)`$; $r = 0.8$ unattenuated).

If we want to test the *block* effects (collapsed across *order*) separately for each presentation time and measure, we get different sensitivities depending on whether we perform correction for multiple comparisons.

Without Bonferroni-correction:

```{r}
otm2_power$emmeans$contrasts

emmeans_power(otm2_power$emmeans$contrasts, alpha_level = 0.1)
```

If we want 80% power for the IAT effect here, we would need $n = 75$ ($N = 150$).
For 95% we need $n = 131$ ($N = 262$).

With Bonferroni-correction

```{r}
emmeans_power(otm2_power$emmeans$contrasts, alpha_level = 0.05)
```

If we want 80% power for the IAT effect here, we would need $n = 96$ ($N = 192$).
For 95% we need $n = 157$ ($N = 314$).

In case presentation time has no effect, we can collapse across presentation times:

```{r}
custom_contrast <- contrast(
  emmeans(
      otm2_power$emmeans$emmeans
      , specs = ~ time * measure * block
  )
  , list(
      eval_block_effect = c(0.5, 0.5, 0, 0, -0.5, -0.5, 0, 0)
      , iat_block_effect = c(0, 0, -0.5, 0.5, 0, 0, 0.5, -0.5)
  )
)

emmeans_power(custom_contrast, alpha_level = 0.1)
```

## Time x Order x Block

```{r eval = TRUE}
knitr::opts_chunk$set(eval = TRUE)
```

-->

## By measure

Because `Superpower` only allows for power analyses, I manually determined the sample size that yields approximately 95% power.
The following assumes that we are not interested in the effect of the counterbalancing factor *order*.

```{r}
otm2_power <- function(f, n, rm_cor = 0.5, alpha = 0.05) {
  contrast_codes <- tibble::tibble(
    time = rep(contr.treatment(2), 4)
    , order = rep(rep(contr.sum(2), each = 2), 2)
    , block = rep(contr.sum(2), each = 4)
    , time_block = time * block
    , time_order = time * order
    , block_order = block * order
    , time_block_order = time * block * order
  )

  mu <- with(
    contrast_codes
    , -order * 0.25*f +
      -block_order * f
  )

  string <- "2w*2b*2b"
  labelnames <- c("block", "1", "2", "order", "p-n", "n-p", "time", "13", "20")

  design_result <- ANOVA_design(
    design = string
    , n = n
    , mu = mu
    , sd = 1
    , r = rm_cor
    , labelnames = labelnames
  )

  ANOVA_exact(
    design_result
    , alpha_level = alpha_level
    , emm = TRUE
    , verbose = FALSE
    , contrast_type = "pairwise"
    , emm_comp = "block | time * order"
  )
}

otm2_n <- 320/4

# 95% power
otm2_pes <- 0.02027
otm2_dz <- pes_to_dz(otm2_pes, otm2_n, n_groups = 4)
rho <- 0.5
f <- dz_to_f(otm2_dz, rho)

otm2_int_power <- otm2_power(f = f, n = otm2_n, rm_cor = rho, alpha = 0.05)

otm1_observed <- readRDS("../../otm1/results/data_processed/otm1_attitudes_cleaned.rds") %>%
    mutate(`IAT score` = scale(IATscore) %>% as.vector)

otm1_observed_iat_plot <- otm1_observed %>%
  ggplot(aes(x = Block, y = `IAT score`, color = ValenceBlock, group = ValenceBlock)) +
    stat_summary(fun.y = mean, geom = "line") +
    stat_summary(fun.data = mean_cl_normal) +
    facet_grid(~ "24/27ms") +
    theme_bw() +
    ggtitle("Results of Experiment 1")

otm2_expected_plot <- otm2_int_power$dataframe %>%
  mutate(
    order = ifelse(order == "order_n-p", "Negative-positive", "Positive-negative")
    , block = ifelse(block == "block_1", "1", "2")
    , time = ifelse(time == "time_13", "13 ms", "20 ms")
  ) %>%
  ggplot(aes(x = block, y = y, color = order, group = order)) +
    stat_summary(fun.y = mean, geom = "line") +
    stat_summary(fun.data = mean_cl_normal) +
    coord_cartesian(ylim = c(-0.6, 0.6)) +
    facet_grid(~ time) +
    guides(color = FALSE) +
    ggtitle("Assumed for power analysis") +
    labs(x = "Block", y = "Attitude") +
    theme_bw()

plot_grid(
  otm1_observed_iat_plot
  , otm2_expected_plot
  , ncol = 2
)
```

```{r}
# 95% power
otm2_pes <- 0.03973071
otm2_dz <- pes_to_dz(otm2_pes, otm2_n, n_groups = 4)
# rho <- 0.5
# f <- dz_to_f(otm2_dz, rho)
# 
# otm2_contrast_power <- otm2_power(f = f, n = otm2_n, rm_cor = rho)$emmeans$contrasts %>%
#   emmeans_power(alpha = 0.05) %>%
#   .[1, ]
# otm2_contrast_power

# 80% power
otm2_pes2 <- 0.0243813
otm2_dz2 <- pes_to_dz(otm2_pes2, otm2_n, n_groups = 4)
# f <- dz_to_f(otm2_dz2, rho)
# 
# otm2_contrast_power2 <- otm2_power(f = f, n = otm2_n, rm_cor = rho)$emmeans$contrasts %>%
#   emmeans_power(alpha = 0.05) %>%
#   .[1, ]
# otm2_contrast_power2
```

```{r}
# 95% power
otm2_dz_between_collapsed <- f_to_dz(dz_to_f(otm2_dz, rho) / sqrt(2), rho)
otm2_pes_between_collapsed <- dz_to_pes(otm2_dz_between_collapsed, otm2_n, n_groups = 4)
# f <- dz_to_f(otm2_dz_between_collapsed, rho)
# 
# otm2_time_collapsed_contrast_power <- otm2_power(f = f, n = otm2_n, rm_cor = rho)$emmeans$emmeans %>%
#   emmeans(specs = ~ block | order) %>%
#   pairs %>%
#   emmeans_power(alpha = 0.05) %>%
#   .[1, ]
# otm2_time_collapsed_contrast_power

# 80% power
otm2_dz2_between_collapsed <- f_to_dz(dz_to_f(otm2_dz2, rho) / sqrt(2), rho)
otm2_pes2_between_collapsed <- dz_to_pes(otm2_dz2_between_collapsed, otm2_n, n_groups = 4)
# f <- dz_to_f(otm2_dz2_between_collapsed, rho)
# 
# otm2_time_collapsed_contrast_power2 <- otm2_power(f = f, n = otm2_n, rm_cor = rho)$emmeans$emmeans %>%
#   emmeans(specs = ~ block | order) %>%
#   pairs %>%
#   emmeans_power(alpha = 0.05) %>%
#   .[1, ]
# otm2_time_collapsed_contrast_power2
```

```{r}
# (Same results apply to two-way and three-way interaction)

# 95% power
otm2_dz_collapsed <- f_to_dz(dz_to_f(otm2_dz, rho) / sqrt(4), rho)
otm2_pes_collapsed <- dz_to_pes(otm2_dz_collapsed, otm2_n, n_groups = 4)
# f <- dz_to_f(otm2_dz_collapsed, rho)
# 
# otm2_collapsed_contrast_power <- otm2_power(f = f, n = otm2_n, rm_cor = rho)$emmeans$emmeans %>%
#   emmeans(specs = ~ block * order) %>%
#   contrast(method = list(collapsed = c(-0.5, 0.5, 0.5, -0.5))) %>%
#   emmeans_power(alpha_level = 0.05)
# otm2_collapsed_contrast_power

# 80% power
otm2_dz2_collapsed <- f_to_dz(dz_to_f(otm2_dz2, rho) / sqrt(4), rho)
otm2_pes2_collapsed <- dz_to_pes(otm2_dz2_collapsed, otm2_n, n_groups = 4)
# f <- dz_to_f(otm2_dz2_collapsed, rho)
# 
# otm2_collapsed_contrast_power2 <- otm2_power(f = f, n = otm2_n, rm_cor = rho)$emmeans$emmeans %>%
#   emmeans(specs = ~ block * order) %>%
#   contrast(method = list(collapsed = c(-0.5, 0.5, 0.5, -0.5))) %>%
#   emmeans_power(alpha_level = 0.05)
# otm2_collapsed_contrast_power2
```

(ref:otm2-pes-plot) Sensitivity curves for OTM2.
The simple contrast is the comparison of the two learning blocks within each *presentation duration*-*valence order* combination (not adjusted for multiple comparisons).
Collapsed sensitivity curves apply to contrasts where data are pooled across either *presentation duration* or *valence order* (valence orders are pooled by reverse coding one order).
$\eta^2_p$-values in the represent effect sizes corresponding to 95% and 80% power.
The grey dashed line at the top represents the smallest *learning block* difference reported by Rydell et al. (2006).
<!-- The power to detect a two-way or three-way interaction is the same as the collapsed once contrast. -->

```{r otm2-pes-plot, fig.width = 7, fig.height = 4, fig.cap = ""}
rydell_power_curve <- tibble(
  rho = seq(-1, 1, length.out = 100)
  , delta = c(
    dz_to_f(rydell_dz, unique(rho)) * 2
    # dz_to_f(mean(heycke_dz), unique(rho)) * 2
  )
  , power = NA
)

otm1_power_curve <- tibble(
  rho = rep(seq(-1, 1, length.out = 100), 2)
  , delta = c(
    dz_to_f(otm1_dz, unique(rho)) * 2
    , dz_to_f(otm1_dz2, unique(rho)) * 2
  )
  , power = rep(c("0.95", "0.80"), each = length(rho)/2)
)

tibble(
  rho = rep(seq(-1, 1, length.out = 100), 2)
  , otm2_delta_simple = c(
    dz_to_f(otm2_dz, unique(rho)) * 2
    , dz_to_f(otm2_dz2, unique(rho)) * 2
  )
  , otm2_delta_between_collapsed = c(
    dz_to_f(otm2_dz_between_collapsed, unique(rho)) * 2
    , dz_to_f(otm2_dz2_between_collapsed, unique(rho)) * 2
  )
  , otm2_delta_collapsed = c(
    dz_to_f(otm2_dz_collapsed, unique(rho)) * 2
    , dz_to_f(otm2_dz2_collapsed, unique(rho)) * 2
  )
  , power = rep(c("0.95", "0.80"), each = length(rho)/2)
) %>% 
  pivot_longer(cols = matches("otm"), names_to = "contrast", values_to = "delta") %>%
  mutate(
    power = relevel(factor(power), "0.95")
    , contrast = gsub("otm2_delta_", "", contrast)
    , contrast = gsub("_", " ", contrast)
    , contrast = factor(contrast, levels = c("simple", "between collapsed", "collapsed"))
  ) %>% 
  ggplot(aes(x = rho, y = delta, color = contrast, group = interaction(power, contrast), linetype = power)) +
  # ggplot(aes(x = rho, y = delta, color = power, group = power)) +
  # geom_line(data = otm1_power_curve, aes(group = power, linetype = power, color = "simple"), alpha = 0.3) +
  geom_line(data = rydell_power_curve, aes(group = power), color = grey(0.8), linetype = "dashed") +
  geom_line() +
  scale_color_viridis_d(
    end = 0.8
    , labels = c(
      bquote("Simple                 "~delta[z] %in% "{"*.(printnum(otm2_dz))*", "*.(printnum(otm2_dz2))*"}")
      , bquote("Collapsed once   "~delta[z] %in% "{"*.(printnum(otm2_dz_between_collapsed))*", "*.(printnum(otm2_dz2_between_collapsed))*"}")
      , bquote("Collapsed twice   "~delta[z] %in% "{"*.(printnum(otm2_dz_collapsed))*", "*.(printnum(otm2_dz2_collapsed))*"}")
    )
  ) +
  ylim(c(0, 1)) +
  labs(
    x = bquote("Correlation between learning blocks ["*rho*"]")
    , y = bquote("Learning block difference [Cohen's"~delta*"]")
    , color = "Contrast"
    , linetype = bquote("Power"~"["*1 - beta*"]")
  ) +
  theme_apa(box = TRUE) + 
  theme(
    panel.grid.major.x = element_line()
    , panel.grid.major.y = element_line()
    , panel.grid.minor.x = element_line()
    , panel.grid.minor.y = element_line()
    # , legend.position = c(0.99, 0.99)
    # , legend.justification = c(1, 1)
    # , legend.background = element_rect(fill = scales::alpha("white", 0.5))
  )
```

### Sensitivty for omnibus-tests

```{r}
# Main effect block and three-way interaction
otm2_power <- function(f, n, rm_cor = 0.5, alpha = 0.05) {
  contrast_codes <- tibble::tibble(
    time = rep(contr.treatment(2), 4)
    , order = rep(rep(contr.sum(2), each = 2), 2)
    , block = rep(contr.sum(2), each = 4)
    , time_block = time * block
    , time_order = time * order
    , block_order = block * order
    , time_block_order = time * block * order
  )

  mu <- with(
    contrast_codes
    , #-order * 0.25*f +
      -block_order * f +
      -block * 0.5 * f +
      time_block_order * f
  )

  string <- "2w*2b*2b"
  labelnames <- c("block", "1", "2", "order", "p-n", "n-p", "time", "13", "20")

  design_result <- ANOVA_design(
    design = string
    , n = n
    , mu = mu
    , sd = 1
    , r = rm_cor
    , labelnames = labelnames
  )

  ANOVA_exact(
    design_result
    , alpha_level = alpha_level
    , emm = FALSE
    , verbose = FALSE
  )
}

otm2_n <- 320/4

# 95% power
otm2_pes <- 0.03973071
otm2_dz <- pes_to_dz(otm2_pes, otm2_n, n_groups = 4)
rho <- 0.5
f <- dz_to_f(otm2_dz, rho)

otm2_time_power <- otm2_power(f = f, n = otm2_n, rm_cor = rho, alpha = 0.05)
otm2_time_power$main_result

# otm2_time_power$dataframe %>%
#   mutate(
#     order = ifelse(order == "order_n-p", "Negative-positive", "Positive-negative")
#     , block = ifelse(block == "block_1", "1", "2")
#     , time = ifelse(time == "time_13", "13 ms", "20 ms")
#   ) %>%
#   ggplot(aes(x = block, y = y, color = order, group = order)) +
#   stat_summary(fun.y = mean, geom = "line") +
#   stat_summary(fun.data = mean_cl_normal) +
#   coord_cartesian(ylim = c(-0.6, 0.6)) +
#   facet_grid(~ time) +
#   guides(color = FALSE) +
#   ggtitle("Assumed for power analysis") +
#   labs(x = "Block", y = "Attitude") +
#   theme_bw()

otm2_dz_block <- f_to_dz(0.5*f, rho = 0.5)
otm2_pes_block <- dz_to_pes(otm2_dz_block, otm2_n, n_groups = 4)

otm2_dz_int <- f_to_dz(f, rho = 0.5)
otm2_pes_int <- dz_to_pes(otm2_dz_int, otm2_n, n_groups = 4)
```